---
title: "Super_Store"
output: html_document
date: '2022-04-13'
---


## Super Store Analysis

Super Store is a small business located in the United States of American. They sell Furniture, Office Supplies and Technology Product and their customers are the mass Consumer, Corporate and Home Offices. My task is to analyze and identify opportunities for Super Store to boost business growth.

First we import all the libraries, this are all the libraries i used

* install.packages(tidyverse)
* install.packages(RColorBrewer)
* install.packages(ggfortify) --convert time series data to a data_Frame
* install.packages(doBy) --To use the order_BY function
* install.packages(plyr)
* install.packages(skimr)
* install.packages(janitor)

# Loading of all the library to be used
```{r}
library(tidyverse)
library(RColorBrewer)
library(ggfortify) 
library(doBy) 
library(plyr)
library(skimr)
library(janitor)
library(scales)
```

## Super Store Exploratory

After installing and loading of my library let see some analysis in our super store dataset

# Loading the dataset

```{r}
super_store<-read.csv("SuperStore.csv")
```
# getting the summarization of all the data in super store
```{r}
skim_without_charts(super_store)
```
#Testing for missing values
```{r}
missing <- sapply(super_store, function(x) sum(is.na(x)))
missing
```
# Checking for the structure of our super stire dataset
```{r}
str(super_store)
```
# Transactions per customer
```{r}
customers<- aggregate(super_store$Customer.Name, by=list(super_store$Customer.Name), FUN=length)
colnames(customers)<-c("Customer","Count")
head(customers)
```
# ploting the graph for trancsaction per customer

```{r, echo=FALSE}
custome = top_n(customers, 20)
ggplot() + geom_bar(data = custome, aes(x = reorder(custome$Customer, -custome$Count), y = custome$Count), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Customers") + ylab("Total Transaction") + ggtitle("Top 20 Customers based on transaction")
```
# Profitability by customer 

```{r}
store_profits <- aggregate(super_store$Profit, by=list(Category=super_store$Customer.Name), FUN=sum)
colnames(store_profits)<-c("Customer","profit")
head(store_profits)
```

# ploting the graph for higest profit per customer

```{r, echo=FALSE}
profits_h =  top_n(store_profits, 20)
ggplot() + geom_bar(data = profits_h, aes(x = reorder(profits_h$Customer, -profits_h$profit), y = profits_h$profit), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Customers") + ylab("Total Transaction") + ggtitle("Top 20 Customers based on profit")
```

##Number of orders from each city

```{r}
citypop <- aggregate(super_store$City, by=list(super_store$City), FUN=length)
colnames(citypop)<-c("City","Count")
head(citypop)
```

```{r, echo=FALSE}
 city_h =  top_n(citypop, 20)
ggplot() + geom_bar(data = city_h, aes(x = reorder(city_h$City, -city_h$Count), y = city_h$Count), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Customers") + ylab("Total Transaction") + ggtitle("Top 20 city based on Transaction")
```

## Total Profit from each city

```{r}
citypor <- aggregate(super_store$Profit, by=list(super_store$City), FUN=sum)
colnames(citypor)<-c("City","profit")
head(citypor)
```

```{r, echo=FALSE}
city_i =  top_n(citypor, 20)
ggplot() + geom_bar(data = city_i, aes(x = reorder(city_i$City, -city_i$profit), y = city_i$profit), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Customers") + ylab("Total Profit") + ggtitle("Top 20 city based on Profit")
```

## Total Sales from each city

```{r}
citysales <- aggregate(super_store$Sales, by=list(super_store$City), FUN=sum)
colnames(citysales)<-c("City","Sales")
head(citysales)
```

```{r, echo=FALSE}
city_s =  top_n(citysales, 20)
ggplot() + geom_bar(data = city_i, aes(x = reorder(city_s$City, -city_s$Sales), y = city_s$Sales), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Customers") + ylab("Total Sales") + ggtitle("Top 20 city based on Sales")
```
## cutomers prefered shipping mode

```{r}
shippment <- aggregate(super_store$Ship.Mode, by=list(super_store$Ship.Mode), FUN=length)
colnames(shippment)<-c("Ship_Mode","Total_Count")
head(shippment)
```

```{r, echo=FALSE}
city_s =  top_n(citysales, 20)
ggplot() + geom_bar(data = shippment, aes(x = reorder(shippment$Ship_Mode, -shippment$Total_Count), y = shippment$Total_Count), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + xlab("Customers") + ylab("Total Sales") + ggtitle("Top 20 city based on Sales")
```

# adding a new column to the data set so as to change the orderdate dataset to Date format

```{r}
new_table <- super_store %>%
  mutate(new_date = parse_date(super_store$Order.Date, "%d-%b-%y"))

view(new_table)
```

# create a daily date
```{r}
dates <- as.Date(new_table$new_date, "%m/%d/%Y")
head(dates)
```

# Extract each days from our dates data
```{r}
days <- weekdays(as.Date(dates, '%Y-%m-%d'))
head(days)
```

# Extracting the year

```{r}
years <- substring(dates, 1,4)
head(years)

Sales_Date <- data.frame(days, super_store$Sales, dates, years)
head(Sales_Date)
```

# plot the graph and change the level of order

```{r echo=FALSE}
Sales_Date$days <- factor(Sales_Date$days, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

ggplot(data = Sales_Date, aes(x=Sales_Date$days, y=super_store$Sales)) + geom_bar(stat = "identity", show.legend = TRUE, fill = "black") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("Total Sales for each year per days") + facet_wrap(~years, ncol = 2) + ylab("Sales in Dollar") + xlab("days")
```

# plot the graph of Yearly sales 

```{r echo=FALSE}
dates1 <- as.Date(new_table$new_date, "%d/%m/%Y")
head(dates1)

saless <- data.frame(dates1, super_store$Sales)
salest <- ts(saless[, -1], frequency = 12, start = c(2014, 03), end = c(2018, 12))
head(salest)

plot(salest, xlab = "years", ylab="sales in dollars", main="Super Store Sales")
```

```{r echo=FALSE}
ggplot() + geom_bar(data = super_store, aes(x = super_store$Category, y = super_store$Sales, fill = Region), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ scale_y_continuous(labels = comma)

ggplot() + geom_bar(data = super_store, aes(x = super_store$Category, y = super_store$Sales), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ scale_y_continuous(labels = comma)
```

# Creating a categories dataframe for both profit and sales
```{r}
catego <- aggregate(super_store$Sales, by=list(super_store$Category), FUN=sum)
colnames(catego)<-c("Categories","Sales")
head(catego)

catepr <- aggregate(super_store$Profit, by=list(super_store$Category), FUN=sum)
colnames(catepr) <- c("Categories", "Profit")
head(catepr)

categor <- data.frame(catego, catepr$Profit)
head(categor)
```

```{r, echo=FALSE}
ggplot() + geom_bar(data = catego, aes(x=categor$Categories, y=categor$Sales), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Total Sales") + xlab("Categories")
ggplot() + geom_bar(data = catego, aes(x=categor$Categories, y=categor$catepr.Profit), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Total Profit") + xlab("Categories")
```

# Creating a segment dataframe for both profot and sales

```{r}
segmen <- aggregate(super_store$Sales, by=list(super_store$Segment), FUN=sum)
colnames(segmen)<-c("Segment","Sales")
head(segmen)

segpro <- aggregate(super_store$Profit, by=list(super_store$Segment), FUN=sum)
colnames(segpro) <- c("Segment", "Profit")
head(segpro)

segmental <- data.frame(segmen, segpro$Profit)
head(segmental)
```

```{r, echo=FALSE}
ggplot() + geom_bar(data = segmental, aes(x=segmental$Segment, y=segmental$Sales), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Total Sales") + xlab("Segment")
ggplot() + geom_bar(data = segmental, aes(x=segmental$Segment, y=segmental$segpro.Profit), stat = "identity") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Total Profit") + xlab("Segment") + scale_y_continuous(labels = comma)
```

## Recommendations 

* Super Store Should focuse on Technology Sub-Category and Phones and Chairs as they are highly selling and profitable. Bundle them with the less profitable product like bookcases and table to offset the losses.

* Seeling Bookcases and Table results in a huge losses, so Super-store has to consider to bundle them with highly selling and profitable product.

* For Home Office Customers, these people might be busy with work and less likely to spend time selecting individual products, so create a home office package with products used for offices.

* For Loss making product, Super store should consider either drop these from our product or we should change suppliers and bargain for cheaper price.

* Consumer and Corporate segment makes up more than 70% of customer base. Let Target them, especially customers from the East and West Region.



